---
title: "Norwegian aid to `r params$selected_country`"
format:
  docx:
    reference-doc: template/country_snapshot_template.docx
    output-file: country_snapshot_output.docx
    number-sections: true
execute: 
  echo: false
  message: false
  warning: false
params:
  selected_country: "Ethiopia"
---

```{r}
#| label: init
#| include: false

# Load packages
source("scripts/setup/load_packages.R")

# Load helper functions used by the report to produce plots, tables, values etc.
purrr::walk(
  list.files("scripts/helpers", pattern = "\\.R$", full.names = TRUE),
  \(x) source(x, local = knitr::knit_global())
)

# Set Norad ggplot theme for plots
noradplot::ggnorad()
```

```{r}
#| label: data-import
#| eval: true
#| include: false

load_final_data <- function() {
  list(
    df_oda_ten     = read_rds(here("data", "final", "df_oda_ten.rds")),
    df_dac_raw     = read_rds(here("data", "final", "df_dac_raw.rds")),
    df_imp_raw     = read_rds(here("data", "final", "df_imp_raw.rds")),
    df_imp_org_raw = read_rds(here("data", "final", "df_imp_org_raw.rds"))
  )
}

final_data <- load_final_data()
```

```{r}
#| label: international-aid
# Run the run_int_total_aid() function with explicit inputs. Returns a named list with plot and summary values.

int_total_aid <- run_int_total_aid(
  data_int_raw = final_data$df_dac_raw,
  selected_country = params$selected_country
)
```

**This country snapshot provides a statistical overview of Norwegian development aid (ODA) to `r params$selected_country`.**

# International aid to `r params$selected_country`

In `r int_total_aid$val_maxyear`, Norway was the `r int_total_aid$val_norway_rank` largest donor of bilateral aid to `r params$selected_country` among members countries of the OECD Development Assistance Committee (DAC). These figures represent the most recent available international data by recipient country.

```{r}
#| label: plot-international-aid
#| fig-dpi: 150
#| fig-width: 5
#| fig-height: 2.4

print(int_total_aid$p_barplot)
```

# Norwegian aid to `r params$selected_country`

```{r}
#| label: total-aid
# Run the run_nor_total_aid() function with explicit inputs. Returns a named list with plot and summary values.
nor_total_aid <- run_nor_total_aid(
  data_nor_bilateral_raw = final_data$df_oda_ten,
  data_nor_multilateral_raw = final_data$df_imp_raw,
  selected_country = params$selected_country
)
```

```{r}
#| label: link-url-aidresults
# Run the run_result_portal_url() function with explicit inputs. Returns a named list with plot and summary values.
result_portal_url <- run_result_portal_url(
  data_nor_bilateral_raw = final_data$df_oda_ten,
  selected_country = params$selected_country
)
```

In `r nor_total_aid$val_bilateral_maxyear`, Norway disbursed NOK `r nor_total_aid$val_bilateral_maxyear_sum` million in bilateral aid to `r params$selected_country`.[^1] In addition, `r params$selected_country` received an estimated share of Norway's core support to multilateral organisations, but such figures for `r nor_total_aid$val_bilateral_maxyear` are not yet available.[^2]

[^1]: The term *bilateral aid* refers to earmarked support to the recipient country provided through both bilateral and multilateral channels. [Aidresults.no](`r result_portal_url`) provides detailed statistics on Norwegian bilateral aid to `r params$selected_country`.

[^2]: Estimated multilateral core support from Norway to `r params$selected_country` has been calculated by the OECD Development Co-operation Directorate (DCD). These estimates, referred to as *imputed multilateral ODA to countries* are based on how multilateral organisations allocate their core resources across recipient countries. Note that aid provided through regional and global programmes beyond multilateral core contributions is not included.

```{r}
#| label: total-aid-plot
#| fig-dpi: 150
#| fig-width: 5
#| fig-height: 2.6

print(nor_total_aid$p_total_aid)
```


## Norwegian bilateral aid in `r nor_total_aid$val_bilateral_maxyear`

```{r}
#| label: bilateral-topics
# Run the run_bilateral_aid() function with explicit inputs and specifying group_var to target_area. Returns a named list with plot and summary values.
bilateral_topics <- run_bilateral_aid(
  data_bilateral = final_data$df_oda_ten,
  selected_country = params$selected_country,
  group_var = "target_area",
  title = "Thematic Distribution"
)
```

In `r bilateral_topics$val_maxyear`, `r bilateral_topics$val_prefix` most of Norway's bilateral aid to `r params$selected_country` was targeting `r bilateral_topics$val_top_labels`.


```{r}
#| label: print-plot-topic
#| fig-height: 2.2
#| fig-width: 5
#| fig-dpi: 150

print(bilateral_topics$p_barplot)
```

```{r}
#| label: partnertype
# Run the run_bilateral_aid() function with explicit inputs and specifying group_var to partner_group_visual. Returns a named list with plot and summary values.

bilateral_partnertype <- run_bilateral_aid(
  data_bilateral = final_data$df_oda_ten,
  selected_country = params$selected_country,
  group_var = "partner_group_visual",
  title = "Partner Distribution"
)
```

In `r bilateral_partnertype$val_maxyear`, `r bilateral_partnertype$val_prefix` Norway's bilateral aid to `r params$selected_country` was channeled through `r bilateral_partnertype$val_top_labels`.

```{r}
#| label: print-plot-partnertype
#| fig-height: 2.2
#| fig-width: 5
#| fig-dpi: 150

print(bilateral_partnertype$p_barplot)
```

```{r}
#| label: partners
# Run the run_partnerlist() function with explicit inputs and specifying data to df_oda_ten for bilateral aid. Returns a named list with table and summary values.

bilateral_partners <- run_partnerlist(
  data = final_data$df_oda_ten,
  partner_col = agreement_partner,
  amount_col = disbursed_mill_nok,
  subtitle_label = "bilateral aid"
)
```

In `r bilateral_partners$val_maxyear`, `r bilateral_partners$val_prefix` the bilateral aid to `r params$selected_country` was directed to `r bilateral_partners$val_top_labels`.

```{r}
#| label: print-table-partner
#| ft.align: "left"

bilateral_partners$table_partnerlist
```

## Estimated Norwegian multilateral core support to `r params$selected_country`

```{r}
#| label: imputed-multi-partners
# Run the run_partnerlist() function with explicit inputs and specifying data to df_imp_org_raw for multilateral aid. Returns a named list with table and summary values.

multilateral_partners <- run_partnerlist(
  data = final_data$df_imp_org_raw,
  partner_col = multilateral_organisasjon,
  amount_col = disbursed_mill_nok,
  subtitle_label = "estimated multilateral core support"
)
```

According to OECD estimates (*imputed multilateral ODA to countries*), `r params$selected_country` received NOK `r nor_total_aid$val_multilateral_maxyear_sum` million of Norwegian core support to multilateral organisations in `r nor_total_aid$val_multilateral_maxyear`.[^3]

[^3]: Due to OECD rounding of values, there may be small discrepancies between the total estimated amount of multilateral core support to `r params$selected_country` in this table and the the figure shown in the table on the first page.

The table below shows the distribution of this estimated core support by multilateral organisation.[^4] In `r multilateral_partners$val_maxyear`, `r bilateral_partners$val_prefix` estimated multilateral core support to `r params$selected_country` was channeled through `r multilateral_partners$val_top_labels`.

[^4]: The table does not include imputed multilateral core support to `r params$selected_country` for all multilateral organisations receiving core support from Norway. However, the estimates include organisations that account for approximately 90 percent of total core contributions from DAC countries to multilateral organisations.

```{r}
#| label: print-table-imputed-org
#| ft.align: "left"

multilateral_partners$table_partnerlist
```

```{r timestamp}
vec_today <- format(Sys.Date(), format="%d. %m. %Y")
```

<br>
<br>
<br>

This country snapshot was generated on `r vec_today` from the [Country Snapshot portal ](http://noradstats.shinyapps.io/country-snapshot/). For questions, please contact Norad's [Seksjon for Statistics and Analysis](mailto:norad-statistikk.og.analyse@norad.no).